---
title: "IS 607 Week 5 Assignment"
author: "Ken Markus"
date: "March 6, 2016"
output: html_document
---
#Assignment – Tidying and Transforming Data

```{r include=FALSE}
#Clear the environment - just make things neater
rm(list=ls())

#Load the necessary libraries
#install.packages("devtools")
#devtools::install_github("hadley/tidyr")

library(tidyr)
library(dplyr)
library(stringr)
```

First we'll load the data and do some basic clean-ups (tidying).
```{r}
#Load the data
theUrl <- "https://raw.githubusercontent.com/kennygfm/IS607/master/NWAirlines.csv"
raw <- read.csv(theUrl, header = TRUE)

#First step is to rename the columns and flatten the data (destinations)
tidy_raw <- raw %>% 
  rename(Airline=X, ontime_flag=X.1) %>% 
  gather("destination","freq",3:7) %>% 
  filter(!is.na(freq)) 
```

Next step is to ensure that Airline name is never blank - I could not figure out how to do this any other way than the following leveraging a for loop. Ideally this would have been possible simply with tidyr or dplyr but I could not conceive of a way
```{r}
for(i in 1:nrow(tidy_raw)) {
  if (i %% 2 == 0) {
    tidy_raw$Airline[i] <- tidy_raw$Airline[i-1]
  }
}

# We rename the observations because manipulating columns with whitespace in their name is onerous
tidy_raw$ontime_flag <- str_replace(tidy_raw$ontime_flag," ","")
```

We should also spread out the rows since ontime and delays are arguably separate variables of the airline-destination observations
```{r}
tidy_raw <- tidy_raw %>% 
  spread(ontime_flag,freq,fill=NA)
```

Let's do a little analysis and find out percent delayed and order by city, this way we can choose the airline with the preferred on-time percentage.
```{r}
mutate(tidy_raw, ontime_perc = round(100* ontime / (ontime + delayed),2)) %>% 
  arrange(destination)
```

Now let's see overall percentages as well
```{r}
tidy_raw %>% 
  group_by(Airline) %>% 
  summarise(ontime=sum(ontime),delayed=sum(delayed)) %>% 
  mutate(ontime_perc = round(100* ontime / (ontime + delayed),2))
```

From this we can see that AM West has a slightly superior on-time percentage. 