---
title: "IS 607 Week 4 Assignment"
author: "Ken Markus"
date: "February 19, 2016"
output: html_document
---
#Question 3a
```{r}
library(stringr)
#Question 3
raw.data <- "555-1239Moe Szyslak(636 555-0113Burns, C. Montgomery555-6542Rev. Timothy Lovejoy555 8904Ned Flanders636-555-3226Simpson, Homer5553642Dr. Julius Hibbert"
name <- unlist(str_extract_all(raw.data, "[[:alpha:]., ]{2,}"))

#a
last_name <- str_replace_all(unlist(str_extract(name,"[[:alpha:]]{2,},")), pattern = ",", replacement ="")
last_name[is.na(last_name)] <- ""

first_name <- str_replace_all(unlist(str_extract(name,",.[[:alpha:]]{1,}.*")), pattern = ",", replacement ="")
first_name[is.na(first_name)] <- ""

one_name <- unlist(str_extract(name,"[[:alpha:]]{1,} [[:alpha:]]{1,}"))
one_name[is.na(one_name)] <- ""

combined <- str_trim(str_c(first_name, last_name, one_name, sep =" "))
```

The above yields the following for combined: 
```{r echo=FALSE}
combined
```

#Question 3b
```{r}

#3b
title_name <- str_detect(name,"Rev.|Dr.")
title_name
```

#3c
```{r}
second_name <- str_detect(combined,"[[:alpha:]]{1,}.* [[:alpha:]]{1,}.* [[:alpha:]]{1,}.*")
second_name
```

#Question 7
```{r}
html_snippet <- "<title>+++BREAKING NEWS+++</title>"

#pull_tag <- str_extract(html,snippet,"<.+>") yields an error or undesired results, the entire string because we are not limiting the result set to the first end tag '>'

#The below works for this example but is too restricting given that many html tages have parameters that include other characters
pull_tag <- str_extract(html_snippet,"<[[:alpha:]]{1,}+>")
pull_tag
```

Let's look at an example where the above would not work due to parameters and fix it.
```{r}
html_snippet <- "<title name='first title'>+++BREAKING NEWS+++</title>"
pull_tag <- str_extract(html_snippet,"<[[:alpha:]]{1,}+>")
pull_tag

#The following will always pull the appropriate first instance of the tag
pull_tag <- pull_tag <- str_extract(html_snippet,"<.+?>")
pull_tag
```

#Question 8
```{r}
bt <- "(5-3)^2=5^2-2*5*3+3^2 conforms to the binomial theorem."
str_extract(bt, "[^0-9=+*()]")
#The above fails because the special characters are not being escaped. A simple fix rectifies the issue
unlist(str_extract_all(bt,"[\\^0-9\\=\\+\\-\\*()]{2,}"))
```

#Question 9
```{r}
gib <- "clcopCow1zmstc0d87wnkig7OvdicpNuggvhryn92Gjuwczi8hqrfpRxs5Aj5dwpn0Tanwo
Uwisdij7Lj8kpf03AT5Idr3coc0bt7yczjatOaootj55t3Nj3ne6c4Sfek.r1w1YwwojigO
d6vrfUrbz2.2bkAnbhzgv4R9i05zEcrop.wAgnb.SqoU65fPa1otfb7wEm24k6t3sR9zqe5
fy89n6Nd5t9kc4fE905gmc4Rgxo5nhDk!gr"
#Let's try all caps first
str_extract_all(gib,"[[:upper:]]")
```

That clearly revealed a message, thanks for calling me a super-nerd!

Let's see if there is another...
```{r}
str_extract_all(gib,"[[:lower:]]")
```

Nope. There may be others, for example where the numbers represent jumps. But for now, let's assume we cracked it...


