---
title: "Markus-Assignment3"
author: "Ken Markus"
date: "February 13, 2016"
output: html_document
---

#Assignment – Combining Data from Multiple Sources

Your task is to create an R dataframe that shows rates of tuberculosis infection by country. You should use the information from the tb database and the attached population.csv file. Your R dataframe should have the following columns: Country, Year, Rate

```{r include=FALSE}

#install.packages("dplyr")
library(plyr)
library(dplyr)
library(knitr)
```

Our first step will be to load the data and clean up the dataframe a bit.

```{r tidy=TRUE, results="hide"}
#Load the TB data
theUrl="https://raw.githubusercontent.com/kennygfm/IS607/master/Week%203/tb.csv"

#Load all of the data into tb_data
tb_data <- read.table(file = theUrl, header = FALSE, sep = ",", col.names = c("country","year","sex","child","adult","elderly"))

#Fix some data elements so we can collapse sex,child,adult,elderly information
tb_data <- tb_data[tb_data$child != -1,]
tb_data <- tb_data[tb_data$adult != -1,]
tb_data <- tb_data[tb_data$elderly != -1,]

#Create a new column that contains a total of TB Cases
tb_data$cases <- tb_data$child + tb_data$adult + tb_data$elderly

#Load all the Population data
theUrl="https://raw.githubusercontent.com/kennygfm/IS607/master/Week%203/population.csv"
population_data <- read.table(file = theUrl, header = TRUE, sep = ",")
```

Now that we have loaded both files into two separate dataframes, we must now merge (join) them. R has a simple merge function to enable this.

```{r tidy=TRUE, results="hide"}
#Merge the two dataframes
tb_combined <- merge(population_data, tb_data, by=c("country","year"))
```

Finally, we must calculate the actual rate as the code below illustrates. Note that we are opting to display rate in per million people given the relatively low frequency.

```{r tidy=TRUE, results="hide"}
#This dataframe can reveal to us the rate by the parameters sex,child,adult,elderly
#Let's create a new dataframe that condenses all of that information, and then calculate the rate
tb_rate <- ddply(tb_combined, .(country,year,population),summarize,total_cases=sum(cases))

#Now, let's calculate the rate, in cases per million
tb_rate$rate <- 1000000 * tb_rate$total_cases / tb_rate$population
tb_rate <- tb_rate[order(tb_rate$country,tb_rate$year),]
```

And the output is...
```{r}
kable(subset(tb_rate, select = c("country","year","rate")), digits=2)
```
